Ecosystem size and complexity as extrinsic drivers of food chain length in branching ecosystems

### 2023-02-26 rmd/manuscript.Rmd
edit: manuscript

The origin of variation in food chain length (FCL), the number of feeding links from the basal species to top predators, has intrigued ecologists for decades due to its importance in regulating top-down effects [@paceTrophicCascadesRevealed1999], energy flow [@wangBiodiversityEcosystemFunctioning2018], and contaminant concentrations in top predators [@kiddEffectsTrophicPosition1998]. Hypotheses regarding controls of FCL can be organized into three broad categories. First, the resource availability hypothesis states that basal productivity controls FCL because imperfect energy conversion through predator-prey interactions restricts energy available to higher trophic levels [@oksanenExploitationEcosystemsGradients1981]. Second, the disturbance hypothesis (or the dynamical stability hypothesis) predicts longer food chains are vulnerable to environmental perturbations (i.e., the dynamical constraint). Thus, habitats with frequent and/or severe disturbance should support fewer trophic levels [@pimmNumberTrophicLevels1977]. Lastly, the ecosystem size hypothesis posits that larger ecosystems should harbor longer food chains [@takimotoEffectsProductivityDisturbance2012; @teruiSpatialDisturbanceSynchrony2019; @postEcosystemSizeDetermines2000].

>>>>

The origin of variation in food chain length (FCL) has intrigued ecologists for decades due to its importance in regulating top-down effects [@paceTrophicCascadesRevealed1999], energy flow [@wangBiodiversityEcosystemFunctioning2018], and contaminant concentrations in top predators [@kiddEffectsTrophicPosition1998]. Hypotheses regarding controls of FCL can be organized into three broad categories. First, the resource availability hypothesis states that basal productivity controls FCL because imperfect energy conversion through predator-prey interactions restricts energy available to higher trophic levels [@oksanenExploitationEcosystemsGradients1981]. Second, the disturbance hypothesis (or the dynamical stability hypothesis) predicts longer food chains are vulnerable to environmental perturbations (i.e., the dynamical constraint). Thus, habitats with frequent and/or severe disturbance should support fewer trophic levels [@pimmNumberTrophicLevels1977]. Lastly, the ecosystem size hypothesis posits that larger ecosystems should harbor longer food chains [@takimotoEffectsProductivityDisturbance2012; @teruiSpatialDisturbanceSynchrony2019; @postEcosystemSizeDetermines2000].
